---
title: 'STAT 595 Midterm 1 Question 2'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Using the Liver Cancer data given on the Beachboard, fit a logistic regression model for the probability of severity based on various predictors (BMI, Age, Time, Markers, Hepatitis, Jaundice). Describe your findings.

```{r}
# Load needed Libraries

library("tidyverse")
library("car")
```


Check data structure and integrity

```{r}
head(Liver_Cancer)

summary(Liver_Cancer)

str(Liver_Cancer)
```

We will start our logistic regression by re-coding the severity variable.High = 0, Low = 1.

```{r}
Liver_Cancer_2 = Liver_Cancer %>% mutate(Severity = ifelse(Severity == "High", 0, 1))

Liver_Cancer_2$Markers = as.factor(Liver_Cancer_2$Markers)
Liver_Cancer_2$Hepatitis = as.factor(Liver_Cancer_2$Hepatitis)
Liver_Cancer_2$Jaundice = as.factor(Liver_Cancer_2$Jaundice)
Liver_Cancer_2$Validation = as.factor(Liver_Cancer_2$Validation)

view(Liver_Cancer_2)
str(Liver_Cancer_2)

```

We will now fit the logisitc regresion model.  Note that R uses logistic regression to predict the P(1) or in our model the P(Low)

```{r}

logit_model = glm(Severity ~ ., data = Liver_Cancer_2,  family = binomial())

summary(logit_model)

```

From the above results we can see that the only significant coeffecient is Age.  All other coeffecients are not significantly different from zero.

We want to check for multicoliniarity

```{r}
vif(logit_model)
```


We will refit the model with just age

```{r}
logit_model_2 = glm(Severity ~ Age, data = Liver_Cancer_2,  family = binomial())

summary(logit_model_2)
```

We want to plot the model to validate our choice of logistic regression

```{r}

newx = seq(10, 85, by = 1)

fm = predict(logit_model_2, newdata = data.frame(Age = newx), type = "response")

plot(newx, fm, type = "l", xlab = "age", ylab = "Probability of Low Liver Cancer", main= "Probabilty of Low Severity", ylim = c(0,1), lwd = 2, col = "red")

grid(NULL, NULL, lty = 6, col = "cornsilk2")

abline(h = 1, lty = 2, lwd = 2)
abline(h = 0, lty = 2, lwd = 2)

```


