---
title: "R Notebook"
output: html_notebook
---

This question should be answered using the Weekly data set, which is part of the ISLR package. This data is similar in nature to the Smarket data from this chapter's lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.

A.	Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?

```{r}
library("tidyverse")
library("ISLR")
library("dplyr")
```

Load the dataset and check structure

```{r}
weekly = ISLR::Weekly

head(weekly)
str(weekly)
```

Start getting numerical and graphical summaries of the data

```{r}

summary(weekly) # Numeric Summary

```

```{r}
# Check correlation of variables
ggcorr(weekly, palette = "RdBu", label = TRUE, label_size = 3, label_round = 2) 

```

```{r}
# Check visual plots of variables
pairs(weekly)
```

B. Use the full data set to perform a logistic regression with
Direction as the response and the five lag variables plus Volume
as predictors. Use the summary function to print the results. Do
any of the predictors appear to be statistically significant? If so,
which ones?

```{r}

weekly_2 = weekly %>% mutate(Direction = ifelse(Direction == "Down", 0, 1))
str(weekly_2)

```

```{r}
weekly_logit_model = glm(Direction ~ . -Year - Today, data = weekly_2,  family = binomial())

summary(weekly_logit_model)
```

C. Compute the confusion matrix and overall fraction of correct
predictions. Explain what the confusion matrix is telling you
about the types of mistakes made by logistic regression.

```{r}

weekly_logit_model_pred = ifelse(predict(weekly_logit_model, type = "response")>0.5, "Up", "Down")


weekly_logit_model_pred = table(predicted = weekly_logit_model_pred, actual = weekly$Direction)

weekly_logit_model_pred

```

D. Now fit the logistic regression model using a training data period
from 1990 to 2008, with Lag2 as the only predictor. Compute the
confusion matrix and the overall fraction of correct predictions
for the held out data (that is, the data from 2009 and 2010).

```{r}

#split the datset
weekly_trn = filter(weekly, Year <=2008)
weekly_test = filter(weekly, Year > 2008)

# Fit model and confusion matrix
Lag2_logit_model = glm(Direction ~ Lag2, data = weekly_trn,  family = binomial())

# Convert to classifications
trn_pred = ifelse(predict(Lag2_logit_model, type = "response")>0.5, "Up", "Down")

# Make Predictions on Training Data
trn_tab = table(predicted = trn_pred, actual_train = weekly_trn$Direction)


#Make Prediction on Test Data
test_pred = ifelse(predict(weekly_test_model, newdata = weekly_test, type = "response")>0.5, "Up", "Down")

test_tab = table(predicted = test_pred, actual_test = weekly_test$Direction)

#View Tables
trn_tab

test_tab
```

E.	Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.

```{r}

# Training with Lag 1

#split the datset
weekly_trn = filter(weekly, Year <=2008)
weekly_test = filter(weekly, Year > 2008)

# Fit model and confusion matrix
Lag1_logit_model = glm(Direction ~ Lag1, data = weekly_trn,  family = "binomial")

# Convert to classifications
trn_pred = ifelse(predict(Lag1_logit_model, type = "response")>0.5, "Up", "Down")

# Make Predictions on Training Data
trn_tab = table(predicted = trn_pred, actual_train = weekly_trn$Direction)

#Make Prediction on Test Data
test_pred = ifelse(predict(Lag1_logit_model, newdata = weekly_test, type = "response")>0.5, "Up", "Down")

test_tab = table(predicted = test_pred, actual_test = weekly_test$Direction)

#View Tables
trn_tab

test_tab

```

```{r}
# Interaction Between Lag1 and Lag2

#split the datset
weekly_trn = filter(weekly, Year <=2008)
weekly_test = filter(weekly, Year > 2008)

# Fit model and confusion matrix
Lag_logit_model = glm(Direction ~ Lag2:Lag1, data = weekly_trn,  family = "binomial")

# Convert to classifications
trn_pred = ifelse(predict(Lag_logit_model, type = "response")>0.5, "Up", "Down")

# Make Predictions on Training Data
trn_tab = table(predicted = trn_pred, actual_train = weekly_trn$Direction)

#Make Prediction on Test Data
test_pred = ifelse(predict(Lag_logit_model, newdata = weekly_test, type = "response")>0.5, "Up", "Down")

test_tab = table(predicted = test_pred, actual_test = weekly_test$Direction)

#View Tables
trn_tab

test_tab

```

```{r}
# Full Model

Direction ~ . -Year - Today

#split the datset
weekly_trn = filter(weekly, Year <=2008)
weekly_test = filter(weekly, Year > 2008)

# Fit model and confusion matrix
Lag_logit_model = glm(Direction ~ . -Year - Today, data = weekly_trn,  family = "binomial")

# Convert to classifications
trn_pred = ifelse(predict(Lag_logit_model, type = "response")>0.5, "Up", "Down")

# Make Predictions on Training Data
trn_tab = table(predicted = trn_pred, actual_train = weekly_trn$Direction)

#Make Prediction on Test Data
test_pred = ifelse(predict(Lag_logit_model, newdata = weekly_test, type = "response")>0.5, "Up", "Down")

test_tab = table(predicted = test_pred, actual_test = weekly_test$Direction)

#View Tables
trn_tab

test_tab
```

