---
title: 'STAT 595 HW 2 Question 2'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Referring to the Body fat data, find the optimal model to predict percent body fat using: Age (years), Weight (lbs), Height (inches), Neck circumference (cm), Chest circumference (cm), Abdomen circumference (cm), Hip circumference (cm), Thigh circumference (cm), Knee circumference (cm), Ankle circumference (cm), Biceps (extended) circumference (cm), Forearm circumference (cm), Wrist circumference (cm)

```{r}

# Load Packages

library("readr")
library("ggplot2")
library("GGally")
library("dplyr")
library("scales")
library("car")
library("janitor")
library("leaps")
library("MASS")
library("Metrics")

# Fix clashes

select = dplyr::select
```


```{r}
# Import Dataset

library(readr)
Body_Fat = read_csv("Body Fat.csv")

head(Body_Fat)

ls(Body_Fat)

# Change variable names to have no spaces or parenthesis

Body_Fat_2 = Body_Fat%>%clean_names()

head(Body_Fat_2)

ls(Body_Fat_2)

```

```{r}
# Descriptive statistics

summary(Body_Fat_2)
```

```{r}
# Check for missing values

sum(is.na(Body_Fat_2))

```

```{r}
# Check how data is categorized

str(Body_Fat_2)
```

Our first step is to see how correlated these variables are

```{r}

ggcorr(Body_Fat_2, palette = "RdBu", label = TRUE, label_size = 3, label_round = 2, hjust = 1)

```

From the above graph and the output we can see that many of them are highly correlated.  This leads us to suspect that there might be multicollinearity between our variables.  We will check the VIF of the linear combinaton

```{r}

# First model fitting for VIF

vif_model_1 = lm(percent_body_fat~ ., data = Body_Fat_2)
summary(vif_model_1)
vif(vif_model_1)

```

We see very few signficant predictors and 3 of our predictors have a VIF higher than 10.  We will start to remove high values of VIF and refit the model to see of we can adjust for multicollinearity


```{r}

# second model fitting for VIF.  Remove Weight_lbs

vif_model_2 = lm(percent_body_fat~ .-weight_lbs, data = Body_Fat_2)
summary(vif_model_2)
vif(vif_model_2)

```

From the second fitting we can see that the number of significant variables has increased and the number of variables with a VIF > 10 is now 2

```{r}

# third model fitting for VIF.  -weight_lbs -abdomen_circumference_cm

vif_model_3 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm, data = Body_Fat_2)
summary(vif_model_3)
vif(vif_model_3)

```

```{r}
# 4th Fitting.  Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm

vif_model_4 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm  , data = Body_Fat_2)
summary(vif_model_4)
vif(vif_model_4)

```

```{r}
# 5th Fitting. Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm

vif_model_5 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm, data = Body_Fat_2)
summary(vif_model_5)
vif(vif_model_5)
```

```{r}
#6th Fitting. Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm

vif_model_6 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm, data = Body_Fat_2)
summary(vif_model_6)
vif(vif_model_6)
```

```{r}
#7th Fitting.  Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm

vif_model_7 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm, data = Body_Fat_2)
summary(vif_model_7)
vif(vif_model_7)
```

```{r}
# 8th Fitting. Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm

vif_model_8 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm, data = Body_Fat_2)
summary(vif_model_8)
vif(vif_model_8)
```

```{r}
#9th Fitting. Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm -biceps_extended_circumference_cm

vif_model_9 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm -biceps_extended_circumference_cm, data = Body_Fat_2)
summary(vif_model_9)
vif(vif_model_9)
```

```{r}
# 10th Fitting. Remove -weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm -biceps_extended_circumference_cm -knee_circumference_cm

vif_model_10 = lm(percent_body_fat~ .-weight_lbs -abdomen_circumference_cm -hip_circumference_cm -thigh_circumference_cm -neck_circumference_cm -chest_circumference_cm -wrist_circumference_cm -biceps_extended_circumference_cm -knee_circumference_cm, data = Body_Fat_2)
summary(vif_model_10)
vif(vif_model_10)
```

```{r}
# Check correlation and calculate R^2, ADJ R^2, and RMSE of model

Lowest_vif = Body_Fat_2 %>% select(-weight_lbs, -abdomen_circumference_cm, -hip_circumference_cm, -thigh_circumference_cm, -neck_circumference_cm, -chest_circumference_cm, -wrist_circumference_cm, -biceps_extended_circumference_cm, -knee_circumference_cm)

ggcorr(Lowest_vif, palette = "RdBu", label = TRUE, hjust = 0.7)

RSS_VIF = c(crossprod(vif_model_10$residuals))
MSE_VIF = RSS_VIF/length(vif_model_10$residuals)
RMSE_VIF = sqrt(MSE_VIF)
R2_VIF = summary(lm(formula = percent_body_fat ~ . - weight_lbs - 
    abdomen_circumference_cm - 
    hip_circumference_cm - thigh_circumference_cm - neck_circumference_cm - 
    chest_circumference_cm - wrist_circumference_cm - 
    biceps_extended_circumference_cm - 
    knee_circumference_cm, data = Body_Fat_2))$r.squared
ajdR2_VIF = summary(lm(formula = percent_body_fat ~ . - weight_lbs - 
    abdomen_circumference_cm - 
    hip_circumference_cm - thigh_circumference_cm - neck_circumference_cm - 
    chest_circumference_cm - wrist_circumference_cm - 
    biceps_extended_circumference_cm - 
    knee_circumference_cm, data = Body_Fat_2))$adj.r.squared
```



We will now perform model selection using CP, R^2 and Adj R^2

Using leaps we will look for the model with the minimum CP, highest Adj R^2 and Highest R^2.  NOTE: R^2 will usually chose the largest possible model and is generally not a good indication.

```{r}

Body_Fat_2 %>% select(c(percent_body_fat, everything()))  # Order variables to make leaps easier

Body_Fat_2.mat = as.matrix(Body_Fat_2)  # Create matrix to work with leaps

cp_model = leaps(x = Body_Fat_2.mat[,2:14], y = Body_Fat_2.mat[,1], names = names(Body_Fat_2)[2:14], method = "Cp")

r2_model = leaps(x = Body_Fat_2.mat[,2:14], y = Body_Fat_2.mat[,1], names = names(Body_Fat_2)[2:14], method = "r2")

adjr2_model = leaps(x = Body_Fat_2.mat[,2:14], y = Body_Fat_2.mat[,1], names = names(Body_Fat_2)[2:14], method = "adjr2")

# Create dataframe of overall CP, R^2, and ADJ R^2
options(warn = -1)
overall_cp_r2_adjr2 = data.frame(cp_model$which, Cp = cp_model$Cp, R2 = r2_model$r2, AdjR2 = adjr2_model$adjr2)

# Select best CP, R^2 and AdjR^2
best_cp = cp_model$which[which((cp_model$Cp == min(cp_model$Cp))),]
best_adjr2 = adjr2_model$which[which((adjr2_model$adjr2 == max(adjr2_model$adjr2))),]
best_r2 = r2_model$which[which((r2_model$r2 == max(r2_model$r2))),]

# Create dataframe of model
options(warn = -1)
model_1 = data.frame(best_cp, best_r2, best_adjr2)

# View Model
model_1

```

```{r}
# Calculate the RMSE, R^2, and ADJR^2 for each model

cp_model =
lm(formula = percent_body_fat ~ . -height_inches -chest_circumference_cm -hip_circumference_cm -knee_circumference_cm -ankle_circumference_cm -biceps_extended_circumference_cm, data = Body_Fat_2)

RSS_CP = c(crossprod(cp_model$residuals))
MSE_CP = RSS_CP/length(cp_model$residuals)
RMSE_CP = sqrt(MSE_CP)
R2_CP = summary(cp_model)$r.squared
ajdR2_CP = summary(cp_model)$adj.r.squared

max_r2_model = lm(formula = percent_body_fat ~ ., data = Body_Fat_2)

RSS_Max_R2 = c(crossprod(max_r2_model$residuals))
MSE_Max_R2 = RSS_Max_R2/length(max_r2_model$residuals)
RMSE_Max_R2 = sqrt(MSE_Max_R2)
R2_Max_R2 = summary(max_r2_model)$r.squared
ajdR2_Max_R2 = summary(max_r2_model)$adj.r.squared

max_adj_r2_model = lm(formula = percent_body_fat ~ . -height_inches -chest_circumference_cm -knee_circumference_cm -ankle_circumference_cm, data = Body_Fat_2)

RSS_Max_Adj_R2 = c(crossprod(max_adj_r2_model$residuals))
MSE_Max_Adj_R2 = RSS_Max_Adj_R2/length(max_adj_r2_model$residuals)
RMSE_Max_Adj_R2 = sqrt(MSE_Max_Adj_R2)
R2_Max_Adj_R2 = summary(max_adj_r2_model)$r.squared
ajdR2_Max_Adj_R2 = summary(max_adj_r2_model)$adj.r.squared

```


We will now do model selection using Forward, Backward, and Stepwise selection with AIC and BIC

```{r}
# Full model for stepwise selection with another for directional

full.model = lm(percent_body_fat ~ ., data = Body_Fat_2)
null.model = lm(percent_body_fat ~ 1, data = Body_Fat_2)

n = nrow(Body_Fat_2)

```


```{r}
# Stepwise AIC

Stepwise_AIC = stepAIC(full.model, direction = "both", trace = FALSE)

summary(Stepwise_AIC)

# Calculate RMSE, R^2, ADJ R^2

RSS_Step_AIC = c(crossprod(Stepwise_AIC$residuals))
MSE_Step_AIC = RSS_Step_AIC/length(Stepwise_AIC$residuals)
RMSE_Step_AIC = sqrt(MSE_Step_AIC)
R2_Step_AIC = summary(Stepwise_AIC)$r.squared
ajdR2_Step_AIC = summary(Stepwise_AIC)$adj.r.squared
```


```{r}
# Stepwise BIC

Stepwise_BIC = stepAIC(full.model, direction = "both", trace = FALSE, k = log(n))

summary(Stepwise_BIC)

# Calculate RMSE, R^2, ADJ R^2

RSS_Step_BIC = c(crossprod(Stepwise_BIC$residuals))
MSE_Step_BIC = RSS_Step_BIC/length(Stepwise_BIC$residuals)
RMSE_Step_BIC = sqrt(MSE_Step_BIC)
R2_Step_BIC = summary(Stepwise_BIC)$r.squared
ajdR2_Step_BIC = summary(Stepwise_BIC)$adj.r.squared
```

```{r}
# Forward AIC

Forward_AIC = stepAIC(null.model,direction="forward",scope=list(upper=full.model,lower=null.model), trace = FALSE)

summary(Forward_AIC)

RSS_Forward_AIC = c(crossprod(Forward_AIC$residuals))
MSE_Forward_AIC = RSS_Forward_AIC/length(Forward_AIC$residuals)
RMSE_Forward_AIC = sqrt(MSE_Forward_AIC)
R2_Forward_AIC = summary(Forward_AIC)$r.squared
ajdR2_Forward_AIC = summary(Forward_AIC)$adj.r.squared
```

```{r}
# Forward BIC

Forward_BIC = stepAIC(null.model,direction="forward",scope=list(upper=full.model,lower=null.model, k=log(n)), trace = FALSE)

summary(Forward_BIC)

RSS_Forward_BIC = c(crossprod(Forward_BIC$residuals))
MSE_Forward_BIC = RSS_Forward_BIC/length(Forward_BIC$residuals)
RMSE_Forward_BIC = sqrt(MSE_Forward_BIC)
R2_Forward_BIC = summary(Forward_BIC)$r.squared
ajdR2_Forward_BIC = summary(Forward_BIC)$adj.r.squared

```

```{r}
# Backward AIC

Backward_AIC = stepAIC(full.model,direction="backward", trace = FALSE)

summary(Backward_AIC)

RSS_Backward_AIC = c(crossprod(Backward_AIC$residuals))
MSE_Backward_AIC = RSS_Backward_AIC/length(Backward_AIC$residuals)
RMSE_Backward_AIC = sqrt(MSE_Backward_AIC)
R2_Backward_AIC = summary(Backward_AIC)$r.squared
ajdR2_Backward_AIC = summary(Backward_AIC)$adj.r.squared
```

```{r}
# Backward BIC

Backward_BIC = stepAIC(full.model,direction="backward", trace = FALSE, k = log(n))

summary(Backward_BIC)

RSS_Backward_BIC = c(crossprod(Backward_BIC$residuals))
MSE_Backward_BIC = RSS_Backward_BIC/length(Backward_BIC$residuals)
RMSE_Backward_BIC = sqrt(MSE_Backward_BIC)
R2_Backward_BIC = summary(Backward_BIC)$r.squared
ajdR2_Backward_BIC = summary(Backward_BIC)$adj.r.squared
```

```{r}
# Model Comparison Dataframe

Model_Comparison = data.frame("R^2 Value" = c(R2_CP, R2_Max_R2, R2_Max_Adj_R2, R2_VIF, R2_Backward_AIC, R2_Backward_BIC, R2_Forward_AIC, R2_Forward_BIC, R2_Step_AIC, R2_Step_BIC), "Adj R^2 Value" = c(ajdR2_CP, ajdR2_Max_R2, ajdR2_Max_Adj_R2, ajdR2_VIF, ajdR2_Backward_AIC, ajdR2_Backward_BIC, ajdR2_Forward_AIC, ajdR2_Forward_BIC, ajdR2_Step_AIC, ajdR2_Step_BIC), "RMSE" = c(RMSE_CP, RMSE_Max_R2, RMSE_Max_Adj_R2, RMSE_VIF, RMSE_Backward_AIC, RMSE_Backward_BIC, RMSE_Forward_AIC, RMSE_Forward_BIC, RMSE_Step_AIC, RMSE_Step_BIC))

rownames(Model_Comparison) = c("CP Model", "Max R^2 Model", "Max.Adj R^2 Model", "VIF Model", "Backward AIC Model", "Backward BIC Model", "Forward AIC Model", "Forward BIC Model", "Stepwise AIC Model", "Stepwise BIC Model")

Model_Comparison
```

```{r}
summary(max_adj_r2_model)
```

Note that 2 of the above predictors are not significant. hip_circumference_cm and biceps_extended_circumference_cm.  We will remove biceps_extended_circumference_cm first since it has the largest P-Vaue and check the new model

```{r}

check = lm(formula = percent_body_fat ~ . - height_inches - chest_circumference_cm - 
    knee_circumference_cm - ankle_circumference_cm -biceps_extended_circumference_cm , data = Body_Fat_2)

summary(check)

RSS_check = c(crossprod(check$residuals))
MSE_check = RSS_check/length(check$residuals)
RMSE_check = sqrt(MSE_check)
R2_checkC = summary(check)$r.squared
ajdR2_check = summary(check)$adj.r.squared

```

This gives us a RMSE of 4.2047

We will remove hip_circumference_cm and check the RMSE

```{r}

check_2 = lm(formula = percent_body_fat ~ . - height_inches - chest_circumference_cm - 
    knee_circumference_cm - ankle_circumference_cm -biceps_extended_circumference_cm -hip_circumference_cm, data = Body_Fat_2)

summary(check_2)

RSS_check_2 = c(crossprod(check_2$residuals))
MSE_check_2 = RSS_check_2/length(check_2$residuals)
RMSE_check_2 = sqrt(MSE_check_2)
R2_check_2 = summary(check_2)$r.squared
ajdR2_check_2 = summary(check_2)$adj.r.squared

```

Note that if we remove the insignificant predictors for the Max_Adj_R2 model we create models that have more insignificant predictors and a lower Adjusted R^2. In conclusion I would recommend that we do not remove the insignificant predictors and keep the model as is.

